
---
title: "Dataset: Seabed litter - Density (Nb. Items/km^2)"
author: Document produced by http://www.marine-analyst.eu
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    number_section: yes
    theme: default
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

```{r setup, include=FALSE}
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(
  eval = TRUE,
	echo = TRUE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	fig.width=5.5,
	out.width = "100%"
)
# clean environment
rm(list=ls())
gc()
```
```{r, include=FALSE, results='hide'}
# Edit the longitude and latitude coordinates to define the geographical area:
minlon=11.3 #minimum longitude
minlat=53.6 #minimum latitude
maxlon=15.5 #maximum longitude
maxlat=55.9 #maximum latitude
wdpaid=paste(minlon,minlat,maxlon,maxlat,sep="_")
Sessionid <- 'test'
wdpaidsplit <- unlist(strsplit(wdpaid, "[_]"))
xmin <- as.numeric(wdpaidsplit[1])
ymin <- as.numeric(wdpaidsplit[2])
xmax <- as.numeric(wdpaidsplit[3])
ymax <- as.numeric(wdpaidsplit[4])
```
```{r, include=FALSE, results='hide'}
source_provider <- "EMODnet Chemistry"
source_provider_url <- "https://www.emodnet.eu"
layer_title<-"Seabed litter - Density (Nb. Items/km^2)"
layer="sl_totalabundance"
wfs_url <- "https://www.ifremer.fr/services/wfs/emodnet_chemistry2?"
wms_url <- "https://www.ifremer.fr/services/wms/emodnet_chemistry2?"
wms_layer="sl_totalabundance"
layer_id<-756
map_legend <- "littercount_by_squarekm"
map_label<-"survey_id"
link_csv<-paste0("./Report-", layer_id, "_", Sessionid, "_", wdpaid, "-csvfile.csv",sep="")
csvfile_name = paste("Report-", layer_id, "_", Sessionid, "_", wdpaid, "-csvfile.csv",sep="")
link_geojson<-paste0("./Report-", layer_id, "_", Sessionid, "_", wdpaid, "-geojsonfile.geojson",sep="")
geojsonfile_name = paste("Report-", layer_id, "_", Sessionid, "_", wdpaid, "-geojsonfile.geojson",sep="")
temp_path<- "."
```

```{r, include=FALSE, results='hide'}
library(rgdal)
library(downloader)
library(ggplot2)
library(mapdata)
library(ggmap)
library(ggrepel)
library(ows4R) # interface for OGC webservices
library(httr) # generic webservice package (build_url)
library(sf) # simple features packages for handling vector GIS data
library(tidyverse) # a suite of packages for data wrangling, transformation, plotting, ...
library(rasterVis)
library(rgeos)
library(sp)
library(raster)
```

# Data information

<TABLE BORDER=0 CELLSPACING=0 CELLPADDING=0 WIDTH=100&#37;><TR><TD WIDTH=100&#37; VALIGN=TOP><DIV ALIGN=JUSTIFY class=><DIV id=standard>This visualization product shows the density of seafloor litter per trawl per year. EMODnet Chemistry included the gathering of marine litter in its 3rd phase. Since the beginning of 2018, data of seafloor litter collected by international fish-trawl surveys have been gathered and processed in the EMODnet Chemistry Marine Litter Database (MLDB). The harmonization of all the data has been the most challenging task considering the heterogeneity of the data sources, sampling protocols (OSPAR and MEDITS protocols) and reference lists used on a European scale. Moreover, within the same protocol, different gear types are deployed during fishing bottom trawl surveys. To calculate the density, the swept area of the trawl is required. The swept area is calculated from the wingspread (which depends on the fishing gear type) and the distance trawled. In the cases were the wingspread and/or the trawled distance are unknown, effective values were calculated from available information, when possible, specifically, the distance: Distance =Haul duration *Ground speed While the calculation of the effective wingspread is based in a general formula (van Hal, 2017) which considers other parameters known of the net, thus introducing some inaccuracies: Wingspread=doordpread*0.18870+5.87280 Doorspread = 14.2 * log(Depth) = 16.72 * log(Warp_length) + 18.49 This method may be improved in the future. The used formula for the swept area and density is: Density (number of items per km^2)= (total number of items)/(Swept area) Percentiles 50, 75 &amp; 95 have been calculated taking into account data from all years. More information is available in the following document: https://www.emodnet-chemistry.eu/repository/20190321_EMODnet_Beach_Seabed_ML_Products.pdf</DIV></DIV></TD></TR></TABLE>

```{r, include=FALSE, results='hide'}
# Script for Wekeo environment
sr=SpatialPolygons(list(Polygons(list(Polygon(cbind(c(xmin, xmin, xmax, xmax),c(ymax, ymin, ymin, ymax)))),"1")))
mpa=SpatialPolygonsDataFrame(sr, data.frame(cbind(1:1), row.names=c("1")))
proj4string(mpa)<-CRS("+proj=longlat +datum=WGS84")
bbox<-paste(xmin,ymin,xmax,ymax,sep=",")
```

```{r, include=FALSE, results='hide'}
# Link to Marine Analyst dataset page
link_marineanalyst <- paste0("http://marine-analyst.eu/dev.py?N=simple&O=",layer_id,"&maxlat=",ymax,"&maxlon=",xmax,"&minlon=",xmin,"&minlat=",ymin)
# Link to open the openlayer page for EMODnet HA
openlayer<-paste0("http://www.marine-analyst.eu/openlayers3/openlayer.py?wms_url=",wms_url,"/wms&wms_layer=",layer,"&bbox=",bbox)
```

## Metadata

Access <A HREF=`r toString(link_marineanalyst)` TARGET=_blank>metadata</A> from landing page<br>

# Geographical extent

## Coordinates

<A HREF=`r toString(link_marineanalyst)` TARGET=_blank>

```{r,echo=FALSE}
print (paste("West-Longitude:",round(xmin,2)))
print (paste("South-Latitude:",round(ymin,2)))
print (paste("East-Longitude:",round(xmax,2)))
print (paste("North-Latitude:",round(ymax,2)))
```

</a>

## Defined area

```{r ,echo=FALSE}
value<-(xmax-xmin)*(ymax-ymin)
if (value > 100) {
      zoom_value<-6
} else if (value > 1) {
      zoom_value<-7
} else {
      zoom_value<-8
}
base<-get_map(location=c(xmin-1,ymin-1,xmax+1,ymax+1), zoom=zoom_value, maptype="terrain-background", source = "stamen")
terrain <- ggmap(base)
map <- terrain + geom_polygon(data=mpa,aes(x=long,y=lat,group=group,fill="mpa"),colour="green",fill="blue",alpha=.1) +
ggtitle("")+xlab("Longitude")+ylab("Latitude")
plot(map)
```

Map tiles by <a href="http://stamen.com">Stamen Design</a>, under <a href="http://creativecommons.org/licenses/by/3.0">CC BY 3.0</a>. Data by <a href="http://openstreetmap.org">OpenStreetMap</a>, under <a href="http://www.openstreetmap.org/copyright">ODbL</a>.

# `r toString(layer_title)`

## Access data

The Web Feature Service (WFS) of the `r toString(source_provider)` portal allows collecting the data:<BR><A HREF=`r toString(wfs_url)`service=WFS&request=GetCapabilities&version=1.1.0 TARGET=_blank>`r toString(wfs_url)`service=WFS&request=GetCapabilities&version=1.1.0</A><BR><BR>**Available labels:**

```{r,echo=FALSE}
# define a function for the DescribeFeatureType request
url <- parse_url(wfs_url)
url$query <- list(service = "WFS",
                  version = "1.0.0",
                  request = "GetCapabilities")
request <- build_url(url)
WFSserver_client <- WFSClient$new(wfs_url, serviceVersion = "1.0.0")
# list of Feature type
WFS_Colnames <- WFSserver_client$
  getCapabilities()$
  findFeatureTypeByName(layer)$
  getDescription() %>%
  map_chr(function(x){x$getName()})
WFS_Colnames
```

```{r, include=FALSE, results='hide'}
url$query <- list(service = "wfs",
                  version = "1.0.0",
                  request = "GetFeature",
                  typename = layer,
                  srsName = "EPSG:4326",
                  outputFormat = "GML3")
request <- build_url(url)
wfs_data <- read_sf(request)
#Transform mpa in Simple feature collection to perform the subsetting because wfs_data contains the whole info - use bbox and tyname are exclusive (Mapserver)
mpaSP <- as(mpa, "SpatialPolygonsDataFrame")
wfs_data<-wfs_data[st_as_sf(mpaSP),]
# get the geometry as lat and lon cols
wfs_data <- wfs_data %>% dplyr::mutate(lat = sf::st_coordinates(.)[,1],lon = sf::st_coordinates(.)[,2])
```

```{r, include=FALSE}
st_write(wfs_data, file.path(temp_path,csvfile_name), layer = csvfile_name, driver = "csv", delete_dsn = TRUE)
st_write(wfs_data, file.path(temp_path,geojsonfile_name), layer = geojsonfile_name, driver = "GeoJSON", delete_dsn = TRUE)
```

<br>**Download data for the defined geographical extent:**<br><TABLE BORDER=1 WIDTH=100%><TR><TD ALIGN=CENTER>Excel file</TD><TD ALIGN=CENTER>Geographic information</TD></TR><TR><TD ALIGN=CENTER><A HREF=`r toString(link_csv)`>**csv**</A></TD><TD ALIGN=CENTER><A HREF=`r toString(link_geojson)` TARGET=_blank>**geojson**</A></TD></TR></TABLE>

## Table

Browse table's columns by using the left and right arrows. Turn the table's pages with help of the previous/next buttons.<br><br>

```{r ,echo=FALSE}
if(nrow(wfs_data) > 0) {
wfs_data
} else {
print("No data available for the defined geographical extent")
}
```

## Map

```{r, echo=FALSE}
if(nrow(wfs_data) > 0) {map <- ggplot() +	borders("worldHires", fill = "gray", colour = "black", xlim = range(xmin,xmax), ylim = range(ymin,ymax), size = .25) +theme(legend.position = "bottom") +    theme(panel.grid.minor.y= element_blank(), panel.grid.minor.x = element_blank()) +    geom_polygon(data=mpa,aes(x=long,y=lat,group=group,fill="mpa"),colour="green",fill="blue",alpha=.1) +	geom_sf() +	geom_point(data = wfs_data, aes(x = lat, y = lon, size =.data[[map_legend]]), fill = "red", color = "red", alpha = .4) +	coord_sf(xlim = c(xmin, xmax),ylim = c(ymin, ymax))+	ggtitle(layer_title)+xlab("Longitude (x)")+ylab("Latitude (y)")map} else {print("No data available for the defined geographical extent")}
```

## Map with id

```{r, echo=FALSE}
if(nrow(wfs_data) > 0) {centroid<- st_centroid(wfs_data)centroid<- cbind(wfs_data, st_coordinates(st_centroid(wfs_data$geometry)))map <- ggplot() +    borders("worldHires", fill = "gray", colour = "black", xlim = range(xmin,xmax), ylim = range(ymin,ymax), size = .25) +theme(legend.position = "bottom") +    theme(panel.grid.minor.y= element_blank(), panel.grid.minor.x = element_blank()) +    geom_polygon(data=mpa,aes(x=long,y=lat,group=group,fill="mpa"),colour="green",fill="blue",alpha=.1) +    geom_sf() +    geom_point(data = wfs_data, aes(x = lat, y = lon, size = .data[[map_legend]]), fill = "red", color = "red", alpha = .4) +    geom_text(data=centroid,aes(x=lat, y=lon, label=.data[[map_label]]), color = "black", fontface = "bold", size=2, hjust= 0, vjust=2, check_overlap = TRUE) +    coord_sf(xlim = c(xmin, xmax),ylim = c(ymin, ymax))+    ggtitle(layer_title)+xlab("Longitude (x)")+ylab("Latitude (y)")map} else {print("No data available for the defined geographical extent")}
```

## Interactive map

Visualise and access data with <A HREF=`r toString(openlayer)` TARGET=_blank>Openlayer (click on the map)<BR>

```{r, include=FALSE}
if(nrow(wfs_data) > 0) {
getWMSmap<-function (wms_layer,xmin,xmax,ymin,ymax)
{
width <- 960
height <- as.integer(width * (ymax-ymin) / (xmax-xmin))
wms_layer<-as.character(wms_layer)
bbox <- paste(xmin, ymin, xmax, ymax, sep = ",")
con<-paste0(wms_url,"/wms?SERVICE=WMS&VERSION=1.1.0&request=GetMap&layers=",wms_layer,"&format=image/jpeg&srs=EPSG:4326&bbox=",bbox,"&height=",height,"&width=",width,"")
        wms <- "img.png"
        wms <- tempfile(wms)
        download(con, wms, quiet = TRUE, mode = "wb")
        img <- brick(wms)
names(img) <- c("img.1", "img.2", "img.3")
img[img$img.1 == 255 & img$img.2 == 255 & img$img.3 == 255] <- NA
wms_basemap_url="http://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv"
wms_basemap_layer="gebco_latest"
con<-paste0(wms_basemap_url,"?SERVICE=WMS&VERSION=1.1.0&request=GetMap&layers=",wms_basemap_layer,"&format=image/png&srs=EPSG:4326&bbox=",bbox,"&height=",height,"&width=",width,"")
        wms <- "img.png"
        wms <- tempfile(wms)
        download(con, wms, quiet = TRUE, mode = "wb")
        basemap <- brick(wms)
names(basemap) <- c("img.1", "img.2", "img.3")
img <- merge(basemap,img)
img@extent@xmin <- xmin
img@extent@ymin <- ymin
img@extent@xmax <- xmax
img@extent@ymax <- ymax
proj4string(img)<-CRS("+proj=longlat +datum=WGS84")
return(img)
}
wms_img<-getWMSmap(wms_layer,xmin,xmax,ymin,ymax)
rggbplot <- function(inRGBRst,npix=NA,scale = 'lin'){
  rgblinstretch <- function(rgbDf){
    maxList <- apply(rgbDf,2,max)
    minList <- apply(rgbDf,2,min)
    temp<-rgbDf
    for(i in c(1:3)){
      temp[,i] <- (temp[,i]-minList[i])/(maxList[i]-minList[i])
    }
    return(temp)
  }
  rgbeqstretch<-function(rgbDf){
    temp<-rgbDf
    for(i in c(1:3)){
      unique <- na.omit(temp[,i])
      if (length(unique>0)){
        ecdf<-ecdf(unique)
        temp[,i] <- apply(temp[,i,drop=FALSE],2,FUN=function(x) ecdf(x))
      }
    }
    return(temp)
  }
      npix <- ncell(inRGBRst)
  x <- sampleRegular(inRGBRst, size=npix, asRaster = TRUE)
  dat <- as.data.frame(x, xy=TRUE)
  colnames(dat)[3:5]<-c('r','g','b')
  if(scale=='lin'){
    dat[,3:5]<- rgblinstretch(dat[,3:5])
  } else if(scale=='stretch'){
    dat[,3:5]<- rgbeqstretch(dat[,3:5])
  }
  p <- ggplot()+ geom_tile(data=dat, aes(x=x, y=y, fill=rgb(r,g,b))) + scale_fill_identity()
}
}
```

```{r, echo=FALSE}
if(nrow(wfs_data) > 0) {
map <- rggbplot(wms_img)+
#borders("worldHires", fill = "gray", colour = "black", xlim = range(xmin,xmax), ylim = range(ymin,ymax), size = .25) +
coord_quickmap(xlim=range(xmin,xmax),ylim=range(ymin,ymax))+
ggtitle(layer_title)+xlab("Longitude")+ylab("Latitude")
plot(map)
} else {
print("No data available for the defined geographical extent")
}
```

</a>

# Litter abundance per year

```{r, echo=FALSE}
if(nrow(wfs_data) > 0) {abundance <- ggplot() +geom_point(data = wfs_data, aes(x = year, y = littercount_by_squarekm, shape=surveyname, group=surveyname ,color=as.character(survey_id)), size=4) +ggtitle("Density per year and location")+xlab("Year")+ylab("Nb. items/km^2")plot(abundance)} else {print("No data available for the defined geographical extent")}
```
